Attack Graph Generation for Microservice Architecture

### 2018-10-21 paper_own_docker/chapters/background.tex
Background updated based on proof-reading feedback

We start by introducing the concept of microservices, their benefits and security implications in Subsection \ref{chap:microservices}. In Subsection \ref{chap:vulnerability_scanners}, we look into vulnerability scanners as tools to scan a single host for vulnerabilities. In Subsection \ref{chap:attack_graphs}, we introduce and formally describe attack graphs as methods to diagnose security weaknesses of a given system composed of multiple hosts.

>>>>

We introduce the concept of microservices, their benefits and security implications in Subsection \ref{chap:microservices}. In Subsection \ref{chap:vulnerability_scanners}, we look into vulnerability scanners as tools to scan a single host for vulnerabilities. In Subsection \ref{chap:attack_graphs}, we introduce and formally describe attack graphs as methods to diagnose the security weaknesses of a given system comprising multiple hosts.

### 2018-10-21 paper_own_docker/chapters/background_attack_graphs.tex
Background updated based on proof-reading feedback

Attack graphs \cite{sheyner2002automated} are a popular way of examining network security weaknesses. They help analysts to analyze a given system and detect its vulnerable parts carefully.  The definition of attack graphs may vary, but it is essentially a directed graph that consists of nodes and edges with various representations.

Seyner et al. define an attack graph as a tuple of states, transitions between the states, initial state and success states. An initial state represents the state from where the attacker starts the attack and through a chain of atomic attacks tries to reach one of the success states \cite{sheyner2002automated}. Ou et al. introduce the notion of a logical attack graph. A logical attack graph is a bipartite directed graph that consists of two kinds of nodes: fact nodes and derivation nodes. Each fact node is labeled with a logical statement in the form of a predicate applied to its arguments, while each derivation node is labeled with an interaction rule that is used for the derivation step. The edges in the graph represent a "depends on" relation \cite{ou2006scalable}. Ingols et al. make a distinction between full, predictive and multiple-prerequisite (MP) attack graphs. Full graph is a directed acyclic graph that consists of nodes that represent hosts and edges that represent vulnerability instances. Predictive attack graphs use the same representation as full attack graphs with the only difference lying in the constraint of when the edges are added to the attack graph. These graphs are generally smaller than full graphs. MP attack is an attack graph with as contentless edges and three node type: state nodes, vulnerability instance nodes and prerequisite nodes \cite{ingols2006practical}.

In this paper, we define an attack graph to be a directed acyclic graph with a set of nodes and edges similar to the full graph representation of Ingols et al. \cite{ingols2006practical}. As an expansion to this model, a node represents a state of a host with its current privilege. An edge represents a successful transition between two such hosts. We can think of an edge as a successful vulnerability exploitation which is initiated from a host with a required privilege to another or the same host with the newly gained privilege as a result of the vulnerability exploitation. To the best of our knowledge, attack graphs are used for networks but not microservices, potentially because there is no tool support.

>>>>

Attack graphs \cite{sheyner2002automated} are a popular way to examine network security weaknesses. They facilitate careful analysis of a given system and detection its vulnerable components. The definition of an attack graph may vary, however, it is essentially a directed graph comprising nodes and edges with various representations.

Seyner et al. defined an attack graph as a tuple of states, transitions between the states, an initial state and success states. An initial state represents the state from which the attacker begins an attack and through a chain of atomic attacks attempts to reach one of the success states \cite{sheyner2002automated}. Ou et al. introduced the notion of a logical attack graph, which is a bipartite directed graph comprising fact and derivation nodes. Each fact node is labeled with a logical statement in the form of a predicate applied to its arguments, while each derivation node is labeled with an interaction rule used in the derivation step. The edges in a logical attack graph represent a "depends on" relation \cite{ou2006scalable}. Ingols et al. made a distinction between full, predictive and multiple-prerequisite (MP) attack graphs. A full graph is a directed acyclic graph comprising of nodes that represent hosts and edges that represent vulnerability instances. Predictive attack graphs use the same representation as full attack graphs with the only difference lying in the constraint of when the edges are added to the attack graph. Note that predictive graphs are generally smaller than full graphs. An MP attack is an attack graph with contentless edges, state nodes, vulnerability instance nodes and prerequisite nodes \cite{ingols2006practical}.

In this paper, we define an attack graph as a directed acyclic graph with a set of nodes and edges similar to the full graph representation proposed by Ingols et al. \cite{ingols2006practical}. As an expansion to this model, a node represents the state of a host with its current privilege, and an edge represents a successful transition between two such hosts. We can consider an edge as a successful vulnerability exploitation initiated from a host with a required privilege to another or the same host with a newly gained privilege as a result of the vulnerability exploitation. To the best of our knowledge, attack graphs have been used for networks but not microservices, potentially because there is currently no existing tool support.

### 2018-10-21 paper_own_docker/chapters/background_microservices.tex
Background updated based on proof-reading feedback

As real-world software grows in size, there is an ever-increasing need to decompose it into an organized structure to promote scaling, reuse, and readability. A software application whose modules cannot be executed independently is called a monolith. Monolithic systems  are characterized by tight coupling, vertical scaling and strong dependence \cite{microservicesfrowler}. Service Oriented Architecture (SOA) addresses these issues by restructuring its elements into components that provide services which are used by other entities through a networking protocol \cite{papazoglou2003service}. However, in a typical SOA, the services are monolithic which gives rise to the concept of microservices  to provide an even more fine-grained task separation \cite{ahmadvand2016requirements}. The novel term "microservices" was first introduced in 2011 at an architectural workshop to propose a common term for the explorations of multiple researchers \cite{dragoni2017microservices, microservicesfrowler}. In the microservices paradigm, multiple services are split into very basic units which are task oriented. According to Dragoni et al. a microservice is a cohesive, independent process interacting via messages. These microservices constitute a distributed architecture called a microservice architecture \cite{dragoni2017microservices}. Microservice architectures benefit us with the advantage of having more heterogeneous technologies, cheaper scaling, resilience, organizational alignment, and composability \cite{newman2015building}. However, they add additional complexity and have a wider attack surface as the need for many services to communicate with each other and third-party software increases \cite{combe2016docker, dragoni2017microservices}. While microservices are an architectural principle, container technology has emerged in cloud computing to provide a lightweight virtualization mechanism. This technology enables microservices to be packaged and orchestrated through the Cloud \cite{pahl2016microservices}.

>>>>

As real-world software increases in size, there is an increasing need to decompose software into an organized structure to promote scalability, reusability, and readability. A software application with modules that cannot be executed independently is referred to as a monolith. Monolithic systems  are characterized by tight coupling, vertical scaling and strong dependence \cite{microservicesfrowler}. The Service Oriented Architecture (SOA) addresses these issues by restructuring its elements into components that provide services that are used by other entities via a networking protocol \cite{papazoglou2003service}. However, in a typical SOA, the services are monolithic which gives rise to the concept of microservices  to provide even more fine-grained task separation \cite{ahmadvand2016requirements}. The term "microservices" was first introduced in 2011 at an architectural workshop as a common term to describe the work of multiple researchers \cite{dragoni2017microservices, microservicesfrowler}. In the microservices paradigm, multiple services are split into very basic task-oriented units. According to Dragoni et al., a microservice is a cohesive, independent process interacting via messages. These microservices constitute a distributed architecture called a microservice architecture \cite{dragoni2017microservices}. Microservice architectures have more heterogeneous technologies, cheaper scaling, resilience, organizational alignment, and composability \cite{newman2015building}. However, they add additional complexity and have a wider attack surface as the need for many services to communicate with each other and third-party software increases \cite{combe2016docker, dragoni2017microservices}. While microservices are an architectural principle, container technology has emerged in cloud computing to provide a lightweight virtualization mechanism. Container technology enables microservices to be packaged and orchestrated through the Cloud \cite{pahl2016microservices}.
